---
title: "Activity Monitoring"
author: "Anushya K"
date: "12/10/2020"
output: html_document
---

# Introduction #

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded here.

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Data Processing ##

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

##load data

data <- read.csv("A:/R/datasets/activity.csv")
```

```{r}
##mean total number of steps taken per day

steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="blue",xlab="Number of Steps")

rmean <- mean(steps_by_day$steps)
rmean

rmedian <- median(steps_by_day$steps)
rmedian


##average daily activity pattern

steps_by_interval <- aggregate(steps ~ interval, data, mean)
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average number of steps per day by interval")

max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval


##imputing missing values

NATotal <- sum(!complete.cases(data))
NATotal


##filling missing values

StepsAverage <- aggregate(steps ~ interval, data = data, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(data)) {
  obs <- data[i, ]
  if (is.na(obs$steps)) {
    steps <- subset(StepsAverage, interval == obs$interval)$steps
  } else {
    steps <- obs$steps
  }
  fillNA <- c(fillNA, steps)
}


##new dataset with imputed missing values

new_activity <- data
new_activity$steps <- fillNA


##histogram

StepsTotalUnion <- aggregate(steps ~ date, data = new_activity, sum, na.rm = TRUE)
hist(StepsTotalUnion$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
#Create Histogram to show difference. 
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "green"), lwd=10)


##calculate mean and median

rmeantotal <- mean(StepsTotalUnion$steps)
rmeantotal

rmediantotal <- median(StepsTotalUnion$steps)
rmediantotal


##comparing with initial values

rmediandiff <- rmediantotal - rmedian
rmediandiff

rmeandiff <- rmeantotal - rmean
rmeandiff


##diff in activity patterns bw weekdays and weekends

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
new_activity$dow = as.factor(ifelse(is.element(weekdays(as.Date(new_activity$date)),weekdays), "Weekday", "Weekend"))
StepsTotalUnion <- aggregate(steps ~ interval + dow, new_activity, mean)
library(lattice)
xyplot(StepsTotalUnion$steps ~ StepsTotalUnion$interval|StepsTotalUnion$dow, main="Average steps per day by interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")


```
